Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-04-21 main.Rmd
diagram changed

\autoref{fig:circular-dow} depicts the linear (a and b) and the corresponding stack (c and d) display of time. Circular granularities day-of-week $C_{B , G}$ and day-of-fortnight $C_{B, H}$ are constructed from linear granularities day ($B$), week ($G$) and fortnight ($H$). The granules that overlap share elements from the underlying time domain in the linear display. The circular granularities $C_{B , G}$ and $C_{B , H}$  <!--= C_{\text{day}, \text{week}}$ representing day-of-week and day-of-fortnight respectively --> consist of repeated pattern of categories $\{0, 1, 2, \dots, 6\}$ and $\{0, 1, 2, \dots, 13\}$ with $P=7$ and $P=14$ respectively. The granules that overlap represents the same category in the stack display. Each circular granularity can use descriptive label mappings, or be simply labelled using the index. <!--Similar to linear granularities, the indexing for each granule of the circular granularity can also be associated with textual representation, called the label.--> Suppose ${L}$ is a label mapping that defines an unique label for each index $l \in \{ 0,1,\dots, (P-1)\}$, then the label mapping $L$ for $C_{B, G}$ can be defined as
$$
  L: \{0,1,2, \dots, 6\} \longmapsto\ \{\text{Sun}, \text{Mon}, \dots, \text{Sat}\}
$$
or
$$
  L: \{0,1,2, \dots, 6\} \longmapsto\ \{\text{Sunday}, \text{Monday}, \dots, \text{Saturday}\}
$$
for example.

@@

```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap=" Circular granularities illustrated through linear (a and b) and stack (c and d) display of time. The linear display is similar to a measuring tape with the measurement marking distributed as linear granularities day, week and fortnight. The stack display gives the intuition of circular granularities by cutting the measuring tape and stacking on top of one another to represent repetitive pattern of categories. The granules that overlap represent same time domain in the linear display and same category in the stack display. Circular granularities day-of-week ($C_{B, G}$) and day-of-fortnight ($C_{B, H}$) have repeating patterns unlike $B$, $G$ or $H$ which gets manifested more clearly through a stack display of time."}
knitr::include_graphics("Figs/circular-example-1.png")
```

@@

```{r aperiodic-example, echo=FALSE, out.width = "100%", fig.cap = "Quasi-circular and aperiodic cyclic granularities illustrated through linear (a) and stack (b) display of time. The linear display shows linear granularities days, weeks, semester weeks, semester week type distributed linearly from past to future. Here a semester lasts for 18.3 weeks, each starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 1.3 weeks. This pattern remains same for all semester and hence $Q_{H, M}$ with $P = 18.3 weeks$ denoting semester week of the semester week type will be a quasi-circular granularity with repeating patterns. $A_{B, M}$ will be an aperiodic cyclic granularity since the placement of the semester within an year varies across years (shown through c)."}

knitr::include_graphics("Figs/quasi.png")

>>>>

\autoref{fig:circular-dow} depicts the linear (a) and the corresponding stack (b) display of time. Circular granularities day-of-week $C_{B , G}$ and day-of-fortnight $C_{B, H}$ are constructed from linear granularities day ($B$), week ($G$) and fortnight ($H$). The granules that overlap share elements from the underlying time domain in the linear display. The circular granularities $C_{B , G}$ and $C_{B , H}$  <!--= C_{\text{day}, \text{week}}$ representing day-of-week and day-of-fortnight respectively --> consist of repeated pattern of categories $\{0, 1, 2, \dots, 6\}$ and $\{0, 1, 2, \dots, 13\}$ with $P=7$ and $P=14$ respectively. The granules that overlap represents the same category in the stack display. Each circular granularity can use descriptive label mappings, or be simply labelled using the index. <!--Similar to linear granularities, the indexing for each granule of the circular granularity can also be associated with textual representation, called the label.--> Suppose ${L}$ is a label mapping that defines an unique label for each index $l \in \{ 0,1,\dots, (P-1)\}$, then the label mapping $L$ for $C_{B, G}$ can be defined as
$$
  L: \{0,1,2, \dots, 6\} \longmapsto\ \{\text{Sun}, \text{Mon}, \dots, \text{Sat}\}
$$
or
$$
  L: \{0,1,2, \dots, 6\} \longmapsto\ \{\text{Sunday}, \text{Monday}, \dots, \text{Saturday}\}
$$
for example.

@@

```{r circular-dow,echo=FALSE,out.width = "100%",fig.cap=" Circular granularities illustrated through linear (a) and stack (b) display of time. The granules that overlap represent same time domain in the linear display and same category in the stack display. Circular granularities day-of-week ($C_{B, G}$) and day-of-fortnight ($C_{B, H}$) have repeating patterns unlike $B$, $G$ or $H$ which gets manifested more clearly through a stack display of time."}
knitr::include_graphics("Figs/circularex.png")
```

@@

```{r aperiodic-example, echo=FALSE, out.width = "100%", fig.cap = "Quasi-circular and aperiodic cyclic granularities illustrated through linear (a) and stack (b) display of time. The linear display shows linear granularities days, weeks, semester weeks, semester week type distributed linearly from past to future. Here a semester lasts for 18.3 weeks, each starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 1.3 weeks. This pattern remains same for all semester and hence $Q_{H, M}$ with $P$ = 18.3 weeks will be a quasi-circular granularity with repeating patterns. $A_{B, M}$ will be an aperiodic cyclic granularity since the placement of the semester within an year varies across years."}

knitr::include_graphics("Figs/quasiex.png")
